Finiteness in Cubical Type Theory

### 2020-05-30 sections/finiteness-predicates.tex
added prose proof of sigma closure

\begin{lemma} \label{split-enum-sigma}
  Split enumerability is closed under \(\Sigma\).
  \begin{equation}
    \frac{
      \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
    }{
      \mathcal{E}!(\Sigma(x : A) , U(x))
    }
  \end{equation}
\end{lemma}

>>>>

With the most basic simple types out of the way, the obvious next choice is the
(non-dependent) sums and products: \(\uplus\) and \(\times\).
Both of these types can be constructed from the \emph{dependent} sum, however,
so that is the type we will prove finite.
From that we can derive a much wider array of finiteness proofs.
\begin{lemma} \label{split-enum-sigma}
  Split enumerability is closed under \(\Sigma\).
  \begin{equation}
    \frac{
      \mathcal{E}!(A) \; \; \; \Pi(x : A) , \mathcal{E}!(U(x))
    }{
      \mathcal{E}!(\Sigma(x : A) , U(x))
    }
  \end{equation}
\end{lemma}
\begin{proof}
  Let \(A\) be a type which is split enumerable, and \(U\) be a type family over
  \(A\) which is split enumerable at every point.
  Formally, we have the following proofs:
  \begin{align}
    \mathcal{E}!_A &: \mathcal{E}!(A) \\
    \mathcal{E}!_U &: \Pi(x : A) , \mathcal{E}!(U(x))
  \end{align}

  Our task is to construct a proof of type:
  \begin{equation}
    \mathcal{E}!(\Sigma(x : A) , U(x))
  \end{equation}
  This proof itself is composed of two components:
  \begin{align}
    \mathit{support} &: \mathbf{List}(\Sigma(x : A) , U(x)) \\
    \mathit{cover} &: \Pi(x : \Sigma(y : A) , U(y)) , x \in \mathit{support}
  \end{align}

  To construct the support list, we apply the function \(\mathcal{E}!_U\) to
  every element in the support list of \(\mathcal{E}!_A\), extract the support
  lists from the resulting finiteness proofs, and concatenate them.

  \begin{agdalisting}
    This pattern of applying a function to each element in a list and
    concatenating the result is of course well-known in functional programming,
    and is in fact the pattern that makes lists a monad.
    While this insight isn't strictly relevant to our work here, it does mean
    the implementation of this function can use Agda's do notation, resulting
    in the following extremely clean implementation:
    \ExecuteMetaData[agda/Cardinality/Finite/SplitEnumerable.tex]{sup-sigma}
  \end{agdalisting}

  To prove that this support list does in fact cover the entirety of the type
  \(\Sigma \; A \; U\), we note that any element of type \(\Sigma \; A \; U\)
  must have a first component in the support list of \(\mathcal{E}!_A\), and its
  second component must be in the result of applying \(\mathcal{E}!_U\) to that
  first element (since that support list contains every element of type
  \(U(x)\)).
  Therefore, the pair itself must be in our constructed support list.
\end{proof}
